---
title: "Mass Shooting Data Report"
author: "Sebastian Liu"
date: "11/22/2020"
output: html_document
---

```{r, include = FALSE}
source("analysis.R")
```

### Summary Information

During the year of **2018**, there were **`r number_of_shootings`** mass
shootings occurred in **the United States**; There were **`r lives_lost`**
people died and **`r ppl_injured`** people injured in mass shootings. Based on
_the total number of deaths and injuries_, **`r most_impacted_city`** was 
_most impacted city_ by shootings, while the state of **`r most_impacted_state`** 
had _the largest total number of deaths and injuries_ among all the states. 
_The worst two mass shootings_ happened on
**`r most_impacted_date[1]` and `r most_impacted_date[2]`**.

### Summary Table

Here is a summary table of mass shootings information **by state** where 
**20 or more people either died or injured** in mass shootings **during 2018**:
```{r, state_info, echo = FALSE, results='asis'}
library(knitr)
knitr::kable(state_info[1:23, ])
```
According to the data, there were **`r nrow(state_info)`** states where mass
shootings caused 20 or more people either died or injured. The worst mass 
shootings happened in **`r state_info[1,1]`** with **`r state_info[1,2]`** 
people lost their lives and **`r state_info[1,3]`** got injured from the 
incidents. Two similar states in terms of mass shooting incidents are
**`r state_info[2,1]`** and **`r state_info[3 ,1]`** with 
**`r state_info[2,2]`** and **`r state_info[3,2]`** people either died or 
injured from shootings. Another two states with large casualties in mass 
shootings are **`r state_info[4,1]`** and **`r state_info[5,1]`** with 
**`r state_info[4,2]`** and **`r state_info[5,2]`** people died from mass 
shootings.

### Description of A Particular Incident

**The worst incident** of the year happened on **`r worst_incident[1,1]`**
at **`r worst_incident[1,4]`** in **`r worst_incident[1,3]`**, 
**`r worst_incident[1,2]`**. The exact location of the incident is 
**(`r worst_incident[1,7]`, `r worst_incident[1,8]`)**. 
**`r worst_incident[1,5]`** people died from the incident, which is the 
**highest** among all the other incidents happened in the U.S. during 2018. 
**`r worst_incident[1,6]`** got injured from the incident, and there were 
**`r worst_incident[1,9]`** people either died or got injured from the incident.
For more information about this incident please click 
**[this link](https://www.nytimes.com/2018/02/14/us/parkland-school-shooting.html)**.

### Interactive Map

```{r, shooting_info, echo = FALSE}
leaflet(data = shooting_info) %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  setView(lng = -100, lat = 40, zoom = 4) %>% 
  addCircles(lat = ~lat,
             lng = ~long,
             stroke = FALSE,
             popup = paste("Date: ", shooting_info$date, "<br>",
                           "City: ", shooting_info$city, "<br>",
                           "Death: ", shooting_info$num_killed, "<br>",
                           "Injured: ", shooting_info$num_injured),
             radius = 40000)
```
As can be seen from the map:

* Most mass shooting incidents happened in east part of the U.S. and in the 
state of California. 

* Mass shooting incidents tend to happen in densely populated areas.

### Plot of My Choice

This graph shows the total casualties of mass shootings **by state** where 
**60 or more people either died or injured** in mass shootings **during 2018**:
```{r, state_plot, echo = FALSE}
ggplotly(state_plot)
```

Point encoding was chosen because _the number of casualties can be quickly_
_identified in this encoding_. This helped me to identify the state with
the highest casualties in mass shootings.

As can be seen from the graph:

* The state of California suffered the most from mass shooting incidents.

* Expect California and Texas, all the other states with casualties more than 60
people are all in the east part of the U.S.