---
title: "Final Project Climate Change"
author: "Bowen"
date: "11/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE, warning=FALSE}
library("knitr")
source("scripts/U.S_CEI_Graph.R")
source("scripts/summary-info-table.R")
source("scripts/monthly_climate_cond_map.R")
source("scripts/temp_precipt_trend_graph.R")
source("scripts/summary_info.R")
```

## Summary Information

___
### Significance

Our project aims to research the effect that climate change has had on the US.
Climate change is something important to address and understand, so we want to
do some data analysis to determine the effect it **already has** on different 
parts of the world and whether it has already begun to impact people's lives. We
used **3** different data sets for our research.

___
### Introduction to Data Sets/Summary Information

#### US Extreme Weather Index

One of the questions that we were interested in was the *number of extreme
weather days that the US had over the last 50 years*. To answer this, we got the
[U.S. Climate Extremes Index (CEI)](https://www.ncdc.noaa.gov/extremes/cei/graph)
, which has a list of the percent of extreme weather days in the United States, 
as well as the number of actual percent of days that are extreme for the year.

There are some important takeaways looking at the data. First, the data set
traces `r cei_summary$years` years, from 1910 to 2019. In these observations,
the year with the most days that were **above the normal temperatures** was 
`r cei_summary$max_high_year`, with `r cei_summary$max_above_norm`% of days
above normal. The year with the highest amount of days **below the normal
temperatures** was `r cei_summary$max_low_year`. In total, for every
`r cei_summary$years` of the observations, the average amount of days that were
**extreme** (low or high) was `r cei_summary$avg_actual_percent`%.

#### Monthly Climate Conditions

Another question that our group was interested in was *How do the average 
temperatures of each state in the United States today compare to the mean 
temperatures of past 50 years or more*. To get the a data set for this, we
went to the **Global Historical Climatology Network-Daily (GHCN-D)** for the
[Monthly Climate Conditions](https://www.ncdc.noaa.gov/cag/divisional/mapping).

The data set **above** has 7 columns for data points within the *contiguous
United States* but we are only using the first four columns. The first and 
second columns are a **ID** for each location, followed by the
**Location Name**. The next column is the **date** of each occurrence, and the 
last column is the **change in temperature** (in Fahrenheit). 

#### Temperature and Precipitation Trends

The last question that our group was interested in was *the trend of average
precipitation and snowfall of each state in the United States over the last
50 years or more* so. This could be found when we compared the **average** 
amount of precipitation per year to each other. To do this, we got the
[Temperature and Precipitation Trends](https://www.ncdc.noaa.gov/cag/national/time-series).

This data set had three columns: the first one has the
**date of the observation**, the next is **average precipitation for the US**
that year, and the last column is the *anomaly*, or
**the teperature from the average**.

___
### Summary Table

##### The following table is the average change in temperature for each state:

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary_table_df <- summary_table()
kable(summary_table_df, col.names = c("State", "Temperature Change"))

min_diff <- min(summary_table_df$change)
max_diff <- max(summary_table_df$change)

name_min <- summary_table_df %>%
  filter(change == min(summary_table_df$change)) %>%
  pull(state)

name_max <- summary_table_df %>%
  filter(change == max(summary_table_df$change)) %>%
  pull(state)

```

This table has all the 48 states in the contiguous US, and their average change
in temperature from the last 50 years.  We can see that the state with the least
change was **`r name_min`** with a **`r min_diff`** increase in temperature. On the
other hand, **`r name_max`** had a **`r max_diff`**, increase, making it the state with
the biggest increase.

___
### Charts

##### The following graph visualizes the Climate Extreme Index:

```{r echo=FALSE, message=FALSE, warning=FALSE}
cei_graph(cei_info)
```

This graph is intended to show the **US Climate Extreme Index** from 1920 - 2019, 
and the relationship between the *approximation* of the percentage of extreme 
weather days with the year. What this graph shows is that the amount of extreme 
weather days was decreasing from 1910 to around 1963, then they began to grow up
to where the current years have the *highest average percentage* of extreme days. 
This difference for modern times can probably be accounted for by global warming, 
and the pollution that the world is getting more and more of. 

##### The following map shows temperature data on the contiguous US:

```{r echo=FALSE, message=FALSE, warning=FALSE}
monthly_climate_cond_map()
```

This is an interactive map that is intended to show the changes in temperature 
by state. From this, we can see that there is no simple correlation between the
state's **location** and the **change in temperature** on the state level. 
Rather, there are sections of the map that have similar temperature changes. The
**Southwest** has a lot of states with massive temp increases, but right next
to it are four states that have some of the least temperature increases.

##### The following graph shows trends of average temperature:

```{r echo=FALSE, message=FALSE, warning=FALSE}
temp_precipt_trend_graph(temp_precip_df)
```

As seen from the title, each of the recorded dates for the year happened in
**October**. This chart is intended to show the relationship between the years
and the average temperature of the US in October. Instead of breaking it down
by **state**, we see the trends when they are grouped as a **nation**. From this
data, the trend is moving upwards, and we can see how average temp is rising.
Another interesting observation is how the anomalies also **increase in both
directions** as the years go by. That shows that there is more variance between
the years, showing that there are more extreme temperatures, *while* the average
continues to climb. 

